\paperTitle

### 2014-05-17 doc/icml2014/background.tex
Minor fixes to background. Added statement about hidden variable relabelling in exclusive views section

For undirected models $\sG$,
let $\sG$ denote a set of cliques, where each clique $\sC \subseteq \bx \cup \bh$ is a subset of nodes.
The joint distribution is given by an exponential family:
  $p_\theta(\bx,\bh) \propto \prod_{\sC \in \sG} \exp(\theta^\top\phi_\sC(\bx_\sC, \bh_\sC))$,
  where $\theta$ is the parameter vector,
  and $\phi_\sC(\bx_\sC, \bh_\sC)$ is the local feature vector
  which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in clique $\sC$.
  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \} \}$ be the neighbors of variable $a$.

@@

Finally, we assume the graphical model is in a canonical form
in which all observed variables are leaves:
\begin{assumption}[Canonical form]
  \label{asm:canonical}
  For each observed variable $x_v$, there exists exactly one $\sC \in \sG$
  such that $\sC = \{ x_v, x_{v'} \}$ for some other node $x_{v'}$.
\end{assumption}
The following lemma shows that this is not a real assumption (see the appendix for the proof):
\begin{lemma}[Reduction to canonical form]
\label{lem:reduction}
Every graphical model can be transformed into canonical
form (via \figureref{reduction}).
There is a one-to-one correspondence between the parameters of the transformed
and original models.
\end{lemma}

@@

\citet{anandkumar12moments} provide an efficient tensor factorization algorithm
for estimating $\Pr(x_v \mid h_i)$:
  %using the method of moments to learn the parameters of a three-view
\begin{theorem}[Tensor factorization]
Let $h_i \in \bh$ be a bottleneck with views $x_{v_1},x_{v_2},x_{v_3}$.
Then there exists an algorithm $\TensorFactorize$ that
returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$.
\end{theorem}

>>>>

For undirected models $\sG$,
let $\sG$ denote a set of cliques, where each clique $\sC \subseteq \bx \cup \bh$ is a subset of nodes.
The joint distribution is given by an exponential family:
  $p_\theta(\bx,\bh) \propto \prod_{\sC \in \sG} \exp(\theta^\top\phi_\sC(\bx_\sC, \bh_\sC))$,
  where $\theta$ is the parameter vector,
  and $\phi_\sC(\bx_\sC, \bh_\sC)$ is the local feature vector
  which only depends on the observed ($\bx_\sC$) and hidden ($\bh_\sC$) variables in clique $\sC$.
  Also define $\sN(a) = \{ b \neq a : \exists \sC \supseteq \{ a, b \} \}$ to be the neighbors of variable $a$.

@@

Finally, we assume the graphical model is in a canonical form
in which all observed variables are leaves:
\begin{assumption}[Canonical form]
  \label{asm:canonical}
  For each observed variable $x_v$, there exists exactly one $\sC \in \sG$
  such that $\sC = \{ x_v, h_{i} \}$ for some hidden node $h_{i}$.
\end{assumption}
The following lemma shows that this is not a real assumption (see the appendix for the proof):
\begin{lemma}[Reduction to canonical form]
\label{lem:reduction}
Every graphical model can be transformed into canonical
form (via \figureref{reduction}).
There is a one-to-one correspondence between the parameters of the transformed
and original models.
\end{lemma}

@@

\citet{anandkumar12moments} provide an efficient tensor factorization algorithm
for estimating $\Pr(x_v \mid h_i)$:
  %using the method of moments to learn the parameters of a three-view
\begin{theorem}[Tensor factorization]
Let $h_i \in \bh$ be a bottleneck with views $x_{v_1},x_{v_2},x_{v_3}$.
Then there exists an algorithm $\TensorFactorize$ that
returns consistent estimates of $\mOpp{v}{i}$ for each $v \in \{v_1,v_2,v_3\}$ up to permutation.
\end{theorem}

### 2014-05-17 doc/icml2014/directed.tex
Minor fixes to background. Added statement about hidden variable relabelling in exclusive views section

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$.
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
the tensor product $\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}

>>>>

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$ up to relabeling of the hidden variables.\footnote{The relabeling for each hidden variables is determined by the permutation in $\TensorFactorize$.}
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
the tensor product $\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}
